---
title: "Autocorrelation function ACF"
output: html_notebook
---
#### Chemical Process Viscosity Readings

```{r}

# Data load


viscosity <- read.csv("viscosity.csv", header= T, sep=";")
head(viscosity)
attach(viscosity)

```
#### Calculating the mean of the readings

```{r}



yHat <- round(mean(Readings),4)
yHat

# just a comment

```
#### Calculating the Autocovariance and Autocorrelation

A time series Fucntion is stationary if the mean of the joint probability distribution of any two observation, y and y(t+k), is the same or any two periods of time t and t+k separated by the same interval K called lag.

If we look on the scatter plot of the data we can see that there is a positive correlation between t and t+1

```{r}

t <- Readings[-length(Readings)]
t_1 <- Readings[-1]

plot(t,t_1, xlab="Reading, Time Period t", ylab="Reading, Time Period t+1",)

```


#### Fuction of the covariance in time Co


The covariance of te time 0 is the variance of the Time Series 

```{r}


# Covariance of the time 0

co <- function(vector){
  
  Co <- c()
  yHat <- mean(vector)
  
  for (i in 1:length(vector)){
  
    Co[i] <- (vector[i]-yHat)*(vector[i]-yHat)
  
  }
  
  Co <- sum(Co)/length(vector)
  Co <- round(Co, 4)
  
  return (Co)
  
}

Co <- co(Readings)
Co
```


```{r}
# Autocovariance lag k = 1

C1 <- c()

for (i in 1:(length(Readings)-1)){
  
  C1[i] <- (Readings[i]-yHat)*(Readings[i+1]-yHat)
  
}
C1 <- sum(C1)/100
C1
```

```{r}

r1 <- C1/Co
r1

```


```{r}
# Autocovariance lag k = 2

C2 <- c()

for (i in 1:(length(Readings)-2)){
  
  C2[i] <- (Readings[i]-yHat)*(Readings[i+2]-yHat)
  
}
C2 <- sum(C2)/100
C2
```
```{r}

# ACF autocorrelation function

ACF <- function(vector,kend = 25){
  
  yHat <- mean(vector)
  Co <- co(vector)
  
  cov_vector <- c()
  rn <- c()
  
  for ( i in 1:kend){
    
    Cn <- c()
    
    for (j in 1:(length(vector)-i)){
  
      Cn[j] <-(vector[j]-yHat)*(vector[j+i]-yHat)
      
    }
    cov_vector[i]<- sum(Cn)/length(vector)
    rn[i] <- cov_vector[i]/Co
  }
  
  #cov_vector < round(cov_vector,4)
  #rn <- round(rn,4)
  result <- data.frame(cov_vector,rn)
  return(result)
  
  
}

```


```{r}

# ACF results

acf <- ACF(Readings)

acf[2]
```

```{r}
library(dplyr)
library(ggplot2)

ggplot(acf, aes(x = seq(1:25), y = rn)) +
  geom_col(position = "identity") + 
  ggtitle("ACF") +
  xlab("Lag, K") + ylab("Autocorrelation, rk")
```

